---
title: "Pipeline step 6 genetic principal components"
---

We include the first 10 genetic PCs since this is generally enough in a European population.

```{r}
library(dplyr) # Run install.packages("tidyverse") to get this
library(tidyr) # Run install.packages("tidyr") to get this
library(ggplot2)
library("reshape") # Run install.packages("reshape") to get this
library(kableExtra) # Run install.packages("kableExtra")

```

Set the date to reference
```{r}

run_date <- "2024-07-29"

```

Set the filepaths for input and output
ENSURE THIS IS A LOCATION SUITABLE FOR STORAGE OF UKB DATA
```{r}

input_path <- # data set storage location

output_path <- # script outputs storage location

```


Load the PCs
```{r}

genetic_PC <- read.csv(paste0(input_path, "ukb_pc20_9-28-22.csv"))

```

```{r}

load(paste0(output_path, "filtered_eid_p2_", run_date, ".RData"))

genetic_PC <- genetic_PC %>%
  filter(genetic_PC$n_eid %in% filtered_eid)

```


Select the first 10
```{r}

genetic_PC <- genetic_PC[,1:11]

```

Rename the columns so they are more readable
```{r}

genetic_PC <- genetic_PC %>%
  rename_with(~ gsub("n_22009_0_", "", .x)) %>%
  rename_with(~ paste0("PC_", .x)) %>%
  # Fix n_eid, it is not a PC
  dplyr::rename("n_eid" = PC_n_eid)

head(genetic_PC)

```

Remove any rows that are missing genetic components

```{r}

print(paste0("Before: ", nrow(genetic_PC)))
n_before <- nrow(genetic_PC)

genetic_PC <- na.omit(genetic_PC)

print(paste0("After: ", nrow(genetic_PC)))
print(paste0((n_before - nrow(genetic_PC)), " participants dropped for missing genetic PCs"))

```

Save the genetic PCs and a list of the IDs which have PCs
```{r}

genetic_PC_eid <- genetic_PC$n_eid

save(genetic_PC_eid, file = paste0(output_path, "genetic_PC_eid_", Sys.Date(), ".RData"))

save(genetic_PC, file = paste0(output_path, "genetic_PC_", Sys.Date(), ".RData"))

```


```{r}

print(paste0("Script last run ", Sys.Date()))

```

